---
title: "Motor_Glu_Paper_Figs"
output: 
  html_document:
    rmdformats::robobook:
      highlight: monochrome
      lightbox: true
      gallery: true
date: "2024-04-23"
---

### Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                   #   out.width="49%",
                   #   out.height="49%",
                      fig.show='hold')
                   #   fig.align='center')
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpmisc)
library(cowplot)

source("~/projects/GluCEST-fMRI/glucest-rsfmri/read_grp_data.R")
#source("/cbica/projects/thalamocortical_development/software/rotate_parcellation/R/perm.sphere.p.R") This package is used for parcel-based spin testing.

#https://www.middleprofessor.com/files/applied-biostatistics_bookdown/_book/adding-covariates-to-a-linear-model

# Markdown help:
# https://bookdown.org/yihui/rmarkdown-cookbook/figures-side.html#fig:figures-side
```

### Set up Environment & Import Data
```{r, message=F, warning=F}
setwd("/Users/pecsok/projects/GluCEST-fMRI/glucest-rsfmri/")

# Set fieldstrength and read in group data:
fieldstrength= "3T"
grp_df <- read_grp_data(fieldstrength,'grp_df_') 
#diag_df = read_grp_data(fieldstrength,'diag_df_')
#view(grp_df)
```

### Set Variables & Choose What to Analyse
```{r, message=F, warning=F}
networks = c("SomMot","Default") #, 
CESTnetworks = c("avgCEST_SomMot","avgCEST_Default") #  "ctCEST_Cont", "ctCEST_Default",, "ctCEST_Limbic",  ,"ctCEST_SalVentAttn"
CESTcts = c("ctCEST_SomMot","ctCEST_Default")

CNB_scores = c("tap_tot")
CNB_valids = c("tap_valid") #, 
diag_scores = c("dx_pscat") # ,"axis1_desc1","axis1_stat1"
demo_scores = c("sex", "age", "race","ethnic","dateDiff")
diag_details = c("axis1_desc1", "axis1_desc2", "axis1_desc3","axis1_desc4","axis1_desc5", "axis1_desc6")

covariate_age = 'age' 
covariate_sex = 'sex'
covariate_ct = 'count'

grp_df$'age' <- as.numeric(grp_df$'age')
grp_df$'avgCEST_SomMot' <- as.numeric(grp_df$'avgCEST_SomMot')
grp_df$'SomMot' <- as.numeric(grp_df$'SomMot')
grp_df$'age' <- as.numeric(grp_df$'age')
grp_df$'dateDiff' <- as.numeric(grp_df$'dateDiff')

grp_df$'sex'[grp_df$'sex'=="1"] <- "M"
grp_df$'sex'[grp_df$'sex'=="2"] <- "F"
grp_df$'diag_group'[grp_df$'hstatus'=="NC"] <- "HC"
grp_df$'diag_group'[grp_df$'hstatus'!="NC"] <- "PSY"
```

### Scrub Data
```{r, message=F, warning=F}
# Remove subjs with outlier scores
grp_df <- grp_df[!(grp_df$BBLID=="20871"),] # SomMot FC < 0.2
grp_df <- grp_df[!(grp_df$BBLID=="80557"),] # tap_tot<75
grp_df <- grp_df[!(grp_df$BBLID=="94703"),] # tap_tot>125
grp_df <- grp_df[!(grp_df$BBLID=="127935"),] # tap_tot>125

print(nrow(grp_df))
```

### QC Analyses
```{r, message=F, warning=F, out.width= "50%"}
# Histograms
vars = c(CESTnetworks, CNB_scores, covariate_age)
for (var in vars) {
  print(var)
  hist(grp_df[[var]],
       main=paste("Histogram of", var,sep=" "),
       xlab=paste(var))
}
```
 
### Linear Models - CEST-age-tapping
```{r dodge-st, message=F, warning=F, out.width= "50%"}
# Initialize an empty list to store ggplot objects
gg_list <- list()

for (CESTnetwork in CESTnetworks) {
  graph_cols <- c(CESTnetwork, 'age', 'sex', 'tap_tot')
  graph_df <- grp_df[complete.cases(grp_df[, graph_cols]), ]
  print(nrow(graph_df))

  # GluCEST vs Tapping Score 
  formula_str <- paste("tap_tot ~", CESTnetwork, "+ age") # + sex
  m1 <- lm(formula_str, data = graph_df)
  print(summary(m1))
  r_squared <- summary(m1)$r.squared
  p_value <- summary(m1)$coefficients[2, 4]
  # Create plot
  gg <- ggplot(graph_df, 
               aes_string(x = CESTnetwork,
                   y = "tap_tot")) + #, color="hstatus"
    geom_point() + #aes(color=treatment)
    geom_smooth(method = "lm",
                formula = y ~ x,
                se = FALSE) +
    annotate("text", x = Inf, y = -Inf, hjust = 1, vjust = 0, 
             label = paste("R-squared =", round(r_squared, 3), "\n", "p-value =", round(p_value, 3)), 
             size = 4) +
    ylab("Tapping Score") +
    xlab(paste(CESTnetwork, "GluCEST")) +
   # theme_pander() +
    theme(legend.position="bottom") 
  # Add ggplot object to the list
  gg_list[[CESTnetwork]] <- gg
}

#plot_grid(gg_list[["avgCEST_SomMot"]], gg_list[["avgCEST_Default"]])

# Print or display ggplot objects
# Print or display ggplot objects
for (gg in gg_list) {
  plot(gg)
}

```


### Linear Models - CEST-age-tapping by diagnosis
```{r dodge-st, message=F, warning=F, out.width= "50%"}
# Initialize an empty list to store ggplot objects
gg_list <- list()

for (CESTnetwork in CESTnetworks) {
  graph_cols <- c(CESTnetwork, 'age', 'sex', 'tap_tot')
  graph_df <- grp_df[complete.cases(grp_df[, graph_cols]), ]
  print(nrow(graph_df))

  # GluCEST vs Tapping Score 
  formula_str <- paste("tap_tot ~", CESTnetwork, "+ age") # + sex
  m1 <- lm(formula_str, data = graph_df)
  print(summary(m1))
  r_squared <- summary(m1)$r.squared
  p_value <- summary(m1)$coefficients[2, 4]
  
  m1_psy <- lm(formula_str, data = graph_df[graph_df$diag_group=='PSY',])
  summary_psy <- summary(m1_psy)
  r_squared_psy <- summary(m1_psy)$r.squared
  p_value_psy <- summary(m1_psy)$coefficients[2, 4]
  print(summary_psy)
  
  m1_hc <- lm(formula_str, data = graph_df[graph_df$diag_group=='HC',])
  summary_hc <- summary(m1_hc)
  r_squared_hc <- summary(m1_hc)$r.squared
  p_value_hc <- summary(m1_hc)$coefficients[2, 4]
  print(summary_hc)
  
    # Create plot
  gg <- ggplot(graph_df, 
               aes_string(x = CESTnetwork,
                   y = "tap_tot",color="diag_group")) + #, 
    geom_point() + #aes(color=treatment)
    geom_smooth(method = "lm",
                formula = y ~ x,
                se = FALSE) +
    annotate("text", x = Inf, y = -Inf, hjust = 1, vjust = 0, 
             label = paste("R-squared =", round(r_squared, 3), "\n", "p-value =", round(p_value, 3)), 
             size = 4) +
    ylab("Tapping Score") +
    xlab(paste(CESTnetwork, "GluCEST")) +
   # theme_pander() +
    theme(legend.position="bottom") 
  # Add ggplot object to the list
  gg_list[[CESTnetwork]] <- gg
}

for (gg in gg_list) {
  plot(gg)
}
```


### Linear Models - FC-age-tapping by diagnosis
```{r dodge-st, message=F, warning=F, out.width= "50%"}
# Initialize an empty list to store ggplot objects
gg_list <- list()

for (network in networks) {
  graph_cols <- c(network, 'age', 'sex', 'tap_tot')
  graph_df <- grp_df[complete.cases(grp_df[, graph_cols]), ]
  print(nrow(graph_df))

  # GluCEST vs Tapping Score 
  formula_str <- paste("tap_tot ~", network, "+ age") # + sex
  m1 <- lm(formula_str, data = graph_df)
  print(summary(m1))
  r_squared <- summary(m1)$r.squared
  p_value <- summary(m1)$coefficients[2, 4]
  
  m1_psy <- lm(formula_str, data = graph_df[graph_df$diag_group=='PSY',])
  summary_psy <- summary(m1_psy)
  r_squared_psy <- summary(m1_psy)$r.squared
  p_value_psy <- summary(m1_psy)$coefficients[2, 4]
  print(summary_psy)
  
  m1_hc <- lm(formula_str, data = graph_df[graph_df$diag_group=='HC',])
  summary_hc <- summary(m1_hc)
  r_squared_hc <- summary(m1_hc)$r.squared
  p_value_hc <- summary(m1_hc)$coefficients[2, 4]
  print(summary_hc)
  
    # Create plot
  gg <- ggplot(graph_df, 
               aes_string(x = network,
                   y = "tap_tot",color="diag_group")) + #, 
    geom_point() + #aes(color=treatment)
    geom_smooth(method = "lm",
                formula = y ~ x,
                se = FALSE) +
    annotate("text", x = Inf, y = -Inf, hjust = 1, vjust = 0, 
             label = paste("R-squared =", round(r_squared, 3), "\n", "p-value =", round(p_value, 3)), 
             size = 4) +
    ylab("Tapping Score") +
    xlab(paste(network, "FC")) +
   # theme_pander() +
    theme(legend.position="bottom") 
  # Add ggplot object to the list
  gg_list[[network]] <- gg
}

for (gg in gg_list) {
  plot(gg)
}
```

### Linear Models - CEST-FC-age-sex-datediff
```{r message=F, warning=F, out.width= "50%"}

# Initialize an empty list to store ggplot objects
gg_list <- list()
sum_list <- list()
#print(grp_df)
for (network in networks) {
  network='SomMot'
  CESTnetwork <- paste("avgCEST_", network, sep = "")
  graph_cols <- c(network, CESTnetwork, 'age', 'dateDiff')
  graph_df <- grp_df[complete.cases(grp_df[, graph_cols]), ]
  graph_df <- grp_df
  # GluCEST vs FC
  formula_str <- paste(CESTnetwork, " ~ dateDiff + age +", network) #sex +
  
  m1 <- lm(formula_str, data = graph_df)
  summary <- summary(m1)
  r_squared <- summary(m1)$r.squared
  p_value <- summary(m1)$coefficients[4, 4]
  # Create plot
  gg <- ggplot(graph_df, 
               aes_string(x = network,
                   y = CESTnetwork)) + #, color=hstatus
    geom_point() + #aes(color=treatment)
    geom_smooth(method = "lm",
                formula = y ~ x,
                se = FALSE) +
    annotate("text", x = Inf, y = -Inf, hjust = 1, vjust = 0, 
             label = paste("R-squared =", round(r_squared, 3), "\n", "p-value =", round(p_value, 3)), 
             size = 4) +
    ylab(CESTnetwork) +
    xlab(paste(network, "FC")) +
   # theme_pander() +
    theme(legend.position="bottom") 
  # Add ggplot object to the list
  gg_list[[CESTnetwork]] <- gg
  sum_list[[CESTnetwork]] <- summary
}

# Print or display ggplot objects
for (gg in gg_list) {
  plot(gg)
}
for (summary in sum_list) {
  print(summary)
}

```


```{r message=F, warning=F, out.width= "50%"}
for (network in networks) {
  CESTnetwork <- paste("avgCEST_", network, sep = "")
  graph_cols <- c(network, CESTnetwork, 'age', 'dateDiff')
  graph_df <- grp_df[complete.cases(grp_df[, graph_cols]), ]
  graph_df <- grp_df
  # GluCEST vs FC
  formula_str <- paste(CESTnetwork, " ~ dateDiff + age +", network) #sex +
  m1 <- lm(formula_str, data = graph_df)
  fitted_models <- graph_df %>% group_by(diag_group) %>% do(model = lm(formula_str, data = graph_df))
  fitted_models$summary
  
}
  
  
```

### Linear Models - CEST-FC-age-sex-datediff by diagnosis
```{r message=F, warning=F, out.width= "50%"}

# Initialize an empty list to store ggplot objects
gg_list <- list()
sum_list <- list()
#print(grp_df)
for (network in networks) {
  CESTnetwork <- paste("avgCEST_", network, sep = "")
  graph_cols <- c(network, CESTnetwork, 'age', 'dateDiff')
  graph_df <- grp_df[complete.cases(grp_df[, graph_cols]), ]
  graph_df <- grp_df
  #view(graph_df)
  print(nrow(grp_df))
  print(nrow(graph_df))
    # GluCEST vs FC
  formula_str <- paste(CESTnetwork, " ~ dateDiff + age +", network) #sex +
  
  m1 <- lm(formula_str, data = graph_df)
  summary <- summary(m1)
  r_squared <- summary(m1)$r.squared
  p_value <- summary(m1)$coefficients[4, 4]
  
  m1_psy <- lm(formula_str, data = graph_df[graph_df$diag_group=='PSY',])
  summary_psy <- summary(m1_psy)
  r_squared_psy <- summary(m1_psy)$r.squared
  p_value_psy <- summary(m1_psy)$coefficients[4, 4]
  
  m1_hc <- lm(formula_str, data = graph_df[graph_df$diag_group=='HC',])
  summary_hc <- summary(m1_hc)
  r_squared_hc <- summary(m1_hc)$r.squared
  p_value_hc <- summary(m1_hc)$coefficients[4, 4]
  
  #print("Summary for Psychosis and " + network)
  print(summary_psy)
  #print("Summary for Healthy and ", network)
  print(summary_hc)
}
```
  # Create plot
  gg <- ggplot(graph_df, 
               aes_string(x = network,
                   y = CESTnetwork,color='diag_group')) + #, 
    geom_point() + #aes(color=treatment)
    geom_smooth(method = "lm",
                formula = y ~ x,
                se = FALSE) +
    annotate("text", x = Inf, y = -Inf, hjust = 1, vjust = 0, 
             label = paste("R-squared =", round(r_squared, 3), "\n", "p-value =", round(p_value, 3)), 
             size = 4) +
    ylab(CESTnetwork) +
    xlab(paste(network, "FC")) +
   # theme_pander() +
    theme(legend.position="bottom") 
  # Add ggplot object to the list
  gg_list[[CESTnetwork]] <- gg
  sum_list[[CESTnetwork]] <- c(summary_psy, summary_hc)
}

# Print or display ggplot objects
for (gg in gg_list) {
  plot(gg)
}
for (summary in sum_list) {
  print(summary)
}

```



